Pop-up cycleways: a peak into post COVID-19 transport futures

### 2020-05-11 conversation.Rmd
Minor tweaks
more references

Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries. As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible. Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.  [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.

Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation...

Evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ]. Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these. This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.

Local and national governments worldwide are still working out how best to respond to these challenges. On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action announced in Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis. Roads can be closed to motor vehicles, or physical barriers can be used to enable [ filtered permeability ], in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles. This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect. Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.

>>>>

Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries. As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible. Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic. [Record bicycle sales](https://www.bbc.co.uk/news/business-52564351) suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.

Active travel has extensive [social and environmental benefits](https://www.bma.org.uk/what-we-do/population-health/preventing-ill-health/steps-to-increase-physical-activity-levels-in-the-uk). Meeting recommended levels of physical activity can help to prevent or manage 20 chronic health conditions, such as cancer, stroke and type 2 diabetes. It can manage or reduce the risk of mental health conditions such as depression and anxiety, and aid with maintaining a healthy weight. Evidence is beginning to emerge of potential links between air pollution and COVID-19 [infection](https://www.ncbi.nlm.nih.gov/pubmed/32315904) and [lethality](https://www.sciencedirect.com/science/article/pii/S0269749120320601), and many health conditions are [associated with poor air quality](https://www.ncbi.nlm.nih.gov/pubmed/30419237). Indeed, across Europe it is estimated that improvements in air quality in recent months have led to [11,000 fewer deaths](https://energyandcleanair.org/wp/wp-content/uploads/2020/04/CREA-Europe-COVID-impacts.pdf) from conditions such as these. This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.

Local and national governments worldwide are still working out how best to respond to these challenges. On Saturday 9th May, British Transport Secretary Grant Shapps announced a [£2 billion investment](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action announced in Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan](https://www.theguardian.com/world/2020/apr/21/milan-seeks-to-prevent-post-crisis-return-of-traffic-pollution) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis. Roads can be closed to motor vehicles, or physical barriers can be used to enable [filtered permeability](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html), in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles. This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect. Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.
